Collective excitations in the 2- and 3-dimensional extended Hubbard model

### 2023-11-24 paper/main.tex

\newcommand{\ddt}{\frac{\mathrm{d}}{\mathrm{d}t}}

@@

Now, the entire Hamiltonian and thus all expectation values only depend on
\begin{equation}
    \hat{\gamma}(\vk) := \sum_{\alpha=1}^D \cos(k_\alpha)\,,
\end{equation}
i.e. for any operator $O$ the following relation holds 
\begin{equation}
    \label{eqn:equal_expecs}
    \langle O_{\vk} \rangle = \langle O_{\vk'} \rangle =: \langle O_{\gamma} \rangle\,,
\end{equation}
if $\hat{\gamma}(\vk) = \hat{\gamma}(\vk')$.
This allows us to replace the momentum sums with energy integrals using the bare density of states
\begin{equation}
    \rho(\gamma) := \frac{1}{N} \sum_{\vk} \delta \left(\gamma - \hat{\gamma} (\vk) \right)\,.
\end{equation}
As an example, consider 
\begin{equation}
    \Delta_\text{SC} = \frac{U}{2N} \int \mathrm{d}\gamma \rho(\gamma) \langle f_{\gamma} \rangle\,.
\end{equation}
We will be using the exact densities of states for the square and the simple cubic lattice found in Ref. \cite{Hanisch97}.
Solving the mean-field equations iteratively, one obtains the groundstate phase diagram at temperature $T=0$ as shown in \autoref{fig:phase_diagram}.
Note, that the $d_{x^2 - y^2}$-superconducting state, which has been confirmed for the square lattice \cite{Micnas88b,Huang13}, cannot be described by our method. 
This is due to our restriction to terms that are proportional to $\hat{\gamma}(\vk)$.
\newline
The CDW-AFM boundary is located at $U = ZV$, where $Z$ is the coordination.
This holds for both the square lattice ($Z=4$) and the simple cubic lattice ($Z=6$) and can be seen by comparing \eqref{eqn:delta_cdw} and \eqref{eqn:delta_afm}:
Crossing the aforementioned line, changes which one of the two prefactors is larger.

@@

Here, the different matrix elements correspond to various Green's functions, depending on the operator basis.
For example, assume the operator in the basis is $f_{\vk}$, then the top left element of $\mathcal{G}$ is the Green's function

@@

>>>>

\newcommand{\ddt}{\frac{\mathrm{d}}{\mathrm{d}t}}
\newcommand{\dgamma}{\mathrm{d}\gamma}

@@

Now, the entire Hamiltonian and thus all expectation values only depend on
\begin{equation}
    \hat{\gamma}(\vk) := \sum_{\alpha=1}^D \cos(k_\alpha)\,,
\end{equation}
i.e. for any operator $O$ the following relation holds 
\begin{equation}
    \label{eqn:equal_expecs}
    \langle O_{\vk} \rangle = \langle O_{\vk'} \rangle =: \langle O_{\gamma} \rangle\,,
\end{equation}
if $\hat{\gamma}(\vk) = \hat{\gamma}(\vk')$.
This allows us to replace the momentum sums with energy integrals using the bare density of states
\begin{equation}
    \rho(\gamma) := \frac{1}{N} \sum_{\vk} \delta \left(\gamma - \hat{\gamma} (\vk) \right)\,.
\end{equation}
As an example, consider 
\begin{equation}
    \Delta_\text{SC} = \frac{U}{2N} \int \dgamma \rho(\gamma) \langle f_{\gamma} \rangle\,.
\end{equation}
We will be using the exact densities of states for the square and the simple cubic lattice found in Ref. \cite{Hanisch97}.
Solving the mean-field equations iteratively, one obtains the groundstate phase diagram at temperature $T=0$ as shown in \autoref{fig:phase_diagram}.
Note, that the $d_{x^2 - y^2}$-superconducting state, which has been confirmed for the square lattice \cite{Micnas88b,Huang13}, cannot be described by our method. 
This is due to our restriction to terms that are proportional to $\hat{\gamma}(\vk)$.
\newline
The CDW-AFM boundary is located at $U = ZV$, where $Z$ is the coordination.
This holds for both the square lattice ($Z=4$) and the simple cubic lattice ($Z=6$) and can be seen by comparing \eqref{eqn:delta_cdw} and \eqref{eqn:delta_afm}:
Crossing the aforementioned line, changes which one of the two prefactors is larger.

@@

Here, the different matrix elements correspond to various Green's functions, depending on the operator basis.
For instance, assume the operator in the basis is $f_{\vk}$, then the top left element of $\mathcal{G}$ is the Green's function

@@

In this section, we will be investigating four different diagonal Green's functions $G_{AA}(\omega + i0^+)$, namely with

\begin{subequations}
    \label{eqn:resolvent_bases}
    \begin{align}
        A_\text{Higgs} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} + f_{vk}^\dagger \right)\,,\\
        A_\text{Phase} &= \frac{i}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} - f_{vk}^\dagger \right)\,,\\
        A_\text{CDW} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} + g_{\vk \down} \right)\,,\\
        A_\text{AFM} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} - g_{\vk \down} \right)\,,
    \end{align}
\end{subequations}

where $N$ is the number of lattice sites in the system. 
Note, that this choice of normalization cancels out with other system size-dependent terms so that our final results do not depend on it at all.
The first operator catches the amplitude mode of the $s$-wave superconducting state, while the second one catches the phase mode \cite{Fan22}.
The remaining two capture the collective behaviour of the CDW and AFM states, respectively.

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview.pdf}
    \caption{Spectral functions with respect to the operators listed in \eqref{eqn:resolvent_bases}.
    The left column shows the results for the square lattice while the right column shows the results for the simple cubic lattice.
    The left edge of the plots is at $-0.05$ rather than $0$ in order to improve the visibility of the phase peak.
    The shaded area marks the two-particle continuum.
    The panels (a), (b), and (c) show the spectral functions in the SC, CDW, and AFM phases, respectively.}
    \label{fig:resolvent_overview}
\end{figure*}

Let us first show and discuss the general form of the spectral functions
\begin{equation}
    \mathcal{A}(\omega) = -\frac{1}{\pi} \Im \left[ \mathcal{G}(\omega + i0^+) \right]
\end{equation}

in various phases and afterward, we will describe the behaviour of the occurring modes as the system approaches a phase transition.
\newline
Firstly, all of the aforementioned operators are (anti-)Hermitian. Therefore, their commutator $[A^\dagger, A]$ vanishes.
Thus, abiding by the sum rule $\int_{-\infty}^\infty \mathcal{A}_{AB} (\omega) \mathrm{d}\omega = \langle [A^\dagger, A] \rangle = 0$ \cite{rickayzen80},
all spectral functions must be symmetrical about $\omega=0$. Therefore, we will only show the $\omega \geq 0$ in our plots.
\newline
Multiple spectral functions in the SC, CDW, and AFM phases are shown in \autoref{fig:resolvent_overview}.
